### 2019-02-05 docs/supplement.Rnw

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\card}[1]{\text{card} \left( #1 \right)}
\makeatletter
\newcommand*{\defeq}{\mathrel{\rlap{%
			\raisebox{0.3ex}{$\m@th\cdot$}}%
		\raisebox{-0.3ex}{$\m@th\cdot$}}%
	=}
\makeatother

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0pt}
\title{Adaptive group-regularized logistic elastic net regression Supplementary material}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}

\begin{document}
\maketitle
\noindent
1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
Amsterdam, The Netherlands\\
2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\

\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes

\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}

\section{Introduction}
This document contains supplementary material to the paper `Adaptive group-regularized logistic elastic net regression'. Section \ref{sec:enetprior} contains extra details on the elastic net prior distribution. Section \ref{sec:objectivefunction} gives the objective function that is maximised in the variational Bayes procedure introduced in Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main Document. The derivations of the variational posterior used in the Main Document are shown in Section \ref{sec:variationaldistributions}. Section \ref{sec:computationalcomplexity} shows how we may reduce the computational complexity of our variational Bayes procedure. Section \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected joint log likelihood that is maximised to estimate the penalty multipliers as explained in Section 3 of the Main Document. The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}. We apply our method to a study on microRNAs in colorectal cancer data \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} . Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.

\section{The elastic net prior}\label{sec:enetprior}
\subsection{Density function}
The density function of $\beta$, drawn from an elastic net prior distribution, is given by:
$$
f(\beta) = g(\lambda_1,\lambda_2) \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\].
$$
The normalizing constant here is calculated as:
$$
g(\lambda_1,\lambda_2) =\left\{ \int_{-\infty}^{\infty} \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta \right\}^{-1}= \frac{\sqrt{\lambda_2}}{2} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1},
$$
with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution functions of $x$, respectively.

\subsection{Expectation and variance}
The expectation and variance of $\beta$ are:
\begin{align*}
\E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
\V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta - \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} - \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
\end{align*}

\subsection{Simulation}
Generating samples from the elastic net prior is not straightforward. \cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$: 
\begin{align*}
\beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} \frac{\tau - 1}{\tau} \right), \\
\tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right),
\end{align*}
where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with an efficient truncated gamma sampler available we may sample from the prior using any standard software package that includes a Gaussian sampling routine. More specifically, we require a a sample from the $\mathcal{TG} \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ distribution. The CDF of this distribution is given by:
$$
F_{T}(\tau) =	
\begin{dcases}
0, & \text{if }\tau < 1 \\
1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}{\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
\end{dcases}
$$

Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for $\tau \geq 1$, the quantile function may be written as a function of the probability $p \in (0,1)$:
$$
Q_{T}(p) =	F^{-1}_{T}(p) = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - p) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2,
$$
where $\Phi^{-1}(p)$ is the quantile function of the standard normal distribution, i.e, the probit function. Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now:
\begin{algorithmic}
	\State{Generate $U \sim \mathcal{U} (0,1)$}
	\State{Set $\tau = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - u) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2$}
	\State{Set $\nu^2= \frac{\tau - 1}{\tau \lambda_2}$}
	\State{Generate $\beta \sim \mathcal{N}(0,\nu^2)$}
	\State{\Return{$\beta$}}
\end{algorithmic}

\section{Objective function}\label{sec:objectivefunction}
The lower bound on the marginal likelihood, or evidence lower bound (ELBO), at iteration $t$ is given by:
\begin{subequations}
	\begin{align}
	\text{ELBO} (Q^{(t)}) & = \E_{Q^{(t)}} [\log \L (\y, \bomega, \bbeta, \btau)] - \E_{Q^{(t)}} [\log Q (\bomega, \bbeta, \btau)] \\
	& = \E_{Q^{(t)}} [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \label{eq:elboomega} \\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\bbeta|\btau)] - \E_{Q^{(t)}} [\log q(\bbeta)] \label{eq:elbobeta}\\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\btau)] - \E_{Q^{(t)}} [\log q(\btau)] \label{eq:elbotau}.
	\end{align}
\end{subequations}

We start with the computation of (\ref{eq:elboomega}). To this end we define:
$$
h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}\frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}\exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
$$
Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, in the following, all expectations and variances are with respect to $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):
\begin{subequations} \label{eq:elboomega2}
	\begin{align}
	\E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\}, \\
	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] \E [\omega_i] \\
	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
	\E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n c_i^2 \E [\omega_i] \\
	& \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
	\end{align}
\end{subequations}
where all mathematical operations on vectors and matrices are element wise. After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
\begin{subequations} \label{eq:elboomega3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
	\end{align}
\end{subequations}

Inspection of the updating equations in the Main Document, learns us that the last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that we are left with just (\ref{eq:elboomega4}). After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
\begin{subequations} \label{eq:elbobeta2}
	\begin{align}
	\E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - \frac{p}{2} \log (2 \pi) \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) (\bmu_j^2 + \bSigma_{jj}) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
	\E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - \frac{p}{2} \log |\bSigma| .
	\end{align}
\end{subequations}
We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):
\begin{subequations} \label{eq:elbobeta3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} [\log q (\bbeta )] \\ 
	& = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + \frac{p}{2} \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
	\end{align}
\end{subequations}

Repeating the same exercise for (\ref{eq:elbotau}) gives:
\begin{subequations} \label{eq:elbotau2}
	\begin{align}
	\E [\log \pi (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{5 p }{2} \log 2 - \frac{p}{2} \log \pi - \frac{p}{2} - \frac{p \lambda_1^2}{8 \lambda_2} \\
	& \,\,\,\,\, - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) - \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} - \frac{1}{2} \sum_{j=1}^p \E [\log ( \psi_j + 1)], \\
	\E [\log q (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{3p}{2} \log 2 -\frac{p}{2} \log \pi -\frac{p}{2} - \frac{\lambda_1}{2 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} \\
	& \,\,\,\,\, + \frac{1}{4} \sum_{j=1}^p \log \chi_j - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j].
	\end{align}
\end{subequations}
Again, we combine the terms in (\ref{eq:elbotau2}) to get:
\begin{subequations} \label{eq:elbotau3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\btau)] - \E_{Q^{(t)}} [\log q (\btau)] 
	\\ & = p \log 2 - \frac{p \lambda_1^2}{8 \lambda_2} - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)} \\
	& \,\,\,\,\, + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log ( \psi_j + 1)].
	\end{align}
\end{subequations}

Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
\begin{align*}
\text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2} \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
& + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
& + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
\end{align*}
Here, proportionality is with respect to the variational and penalty parameters.

\section{The P\'{o}lya-Gamma parametrisation}\label{sec:polyagamma}
In Section 3 of the Main Document latent P\'{o}lya-Gamma distributed variables are introduced. Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma variables are introduced into the logistic regression model under a Gaussian prior. Here, we explain how introduction of these latent variables simplifies the variational Bayes calculations.

In \cite{polson_bayesian_2013}, the central identity for this parametrisation is given in their Theorem 1:
\begin{equation}\label{eq:intid}
\frac{\exp(\psi)^{a}}{[1 + \exp(\psi)]^{b}} = 2^{-b} \exp(\kappa \psi) \int_0^{\infty} \exp(-\omega \psi^2 /2) \pi(\omega) d \omega,
\end{equation}
where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed. This holds for all $b > 0$ and $a \in \mathbb{R}$. After setting $a=y_i$, $b=m_i$, and $\psi=\x_i \tr \bbeta$, we may consider the right-hand side of (\ref{eq:intid}) as an un-normalised joint density in $(\bbeta, \omega)$. The likelihood contribution of observation $i$ for $\bbeta$ conditional on $\omega_i$ may now be written as:
$$
\L_i(\bbeta | \omega_i) \propto \exp[\kappa_i \x_i \tr \bbeta - \omega_i (\x_i \tr \bbeta)^2/2].
$$

This allows the conditional posterior of $\bbeta$ to be written as:
$$
p(\bbeta | \bomega, \y) \propto \pi(\bbeta) \prod_{i=1}^n \L_i(\bbeta | \omega_i) = \pi(\bbeta) \exp \[-\frac{1}{2}(\bOmega^{-1} \bkappa - \X \bbeta) \tr \bOmega (\bOmega^{-1} \bkappa - \X \bbeta)\],
$$
where $\bOmega = \diag (\omega_i)$, for $i=1, \dots, n$. Consequently, with $\pi(\bbeta)$ a Gaussian prior, $p(\bbeta | \bomega, \y)$ is also in the Gaussian family.

Furthermore, by treating the integrand in (\ref{eq:intid}) as an un-normalized density over $(\psi, \phi)$, we find the distribution of $\omega$ conditional on $\bbeta$,
$$
p(\omega | \psi) = \frac{\exp(-\omega \psi^2/2) \pi(\omega)}{\int_0^{\infty} \exp(-\omega \psi^2/2) \pi(\omega) d\omega},
$$
distributed as $\omega | \psi \sim \mathcal{PG}(b,\psi)$ \cite[]{polson_bayesian_2013}, with expectation $\E[ \omega ] = \frac{b}{2\psi} \tanh (\psi/2)$. We substitute again $b=m_i$ and $\psi=\x_i \tr \bbeta$, and set the prior $\pi(\bbeta)$ to a Gaussian with mean $\m$ and covariance $\Vm$ to arrive at the following conditional posteriors:
\begin{subequations}\label{eq:gibbssampler}
	\begin{align}
	\omega_i | \bbeta & \sim \mathcal{PG}(m_i, \x_i \tr \bbeta) \\
	\bbeta | \y, \bomega & \sim \N(\bmu, \bSigma),
	\end{align}
\end{subequations}
where
\begin{align*}
\bSigma & = (\X \tr \bOmega \X + \Vm^{-1})^{-1}, \\
\bmu & = \Vm (\X \tr \bkappa + \Vm^{-1} \m),
\end{align*}
with $\bkappa = \begin{bmatrix} \kappa_1 \dots \kappa_n \end{bmatrix} \tr$. Equations (\ref{eq:gibbssampler}) constitute the Gibbs sampler used in \cite{polson_bayesian_2013}. However, from the corresponding variational Bayes derivations in Section \ref{sec:variationaldistributions}, it is obvious that the parametrisation above is also useful in the variational Bayes approximation to the posterior distribution.

\section{Variational posterior}\label{sec:variationaldistributions}
The optimal distributions in our variational Bayes implementation are:
\begin{align}
q^*_{\bbeta} (\bbeta) & {\propto} \exp \{ \E_{\bomega, \btau} [\log p (\bbeta, \bomega, \btau, \y)]\} {\propto} \exp \{\log \L (\y ; \bbeta) + \E_{\bomega} [\log \pi (\bomega | \bbeta)] + \E_{\btau} [ \log \pi (\bbeta | \btau)] \}, \label{eq:qbeta} \\
q^*_{\bomega} (\bomega) & {\propto} \exp \{ \E_{\bbeta, \btau} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi (\bomega | \bbeta)] \},\label{eq:qomega} \\
q^*_{\tau} (\btau) & {\propto} \exp \{ \E_{\bbeta, \bomega} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi ( \bbeta | \btau)] + \log \pi (\btau) \}. \label{eq:qtau}
\end{align}

For convenience we work on the log-scale and occasionally do a change of variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three terms on the right-hand side of (\ref{eq:qbeta}):
\begin{align*}
\log \L (\y ; \bbeta) & = \sum_{i=1}^n \log f (y_i ; \bbeta) \propto \sum_{i=1}^n \log \left\{ \frac{\exp(\x_i \tr \bbeta)^{y_i}}{[1 + \exp(\x_i \tr \bbeta)]^{m_i}} \right\} \\
& = \y \tr \X \bbeta - \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)], \\
\E_{\bomega} [\log \pi (\bomega | \bbeta)] & \propto \sum_{i=1}^n \E_{\omega_i} [\log \pi (\omega_i | \bbeta)] \\ & \propto \sum_{i=1}^n \E_{\omega_i} \( \log \left\{ \cosh\( \frac{\x_i \tr \bbeta}{2}\)^{m_i} \exp \[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i \]\right\} \) \\
& \propto \sum_{i=1}^n m_i \log \cosh \( \frac{\x_i \tr \bbeta }{2} \) - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta \\ & \propto \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)] - \frac{1}{2} \m \tr \X \bbeta - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta, \\
\E_{\btau} [ \log \pi (\bbeta | \btau)] & = \sum_{j=1}^p \E_{\tau_j} [ \log \pi (\beta_j | \tau_j)] \propto \frac{\lambda_2}{2} \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} \E \( \frac{\tau_j}{\tau_j - 1} \) \\
& = \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} [1 + \E ( \psi_j^{-1} )] = - \frac{1}{2} \lambda_2 \bbeta \tr \bLambda' (\I + \Z) \bbeta,
\end{align*}
where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega = \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' = \diag (\lambda'_{g(j)})$. 
Since the difficult $\sum_{i=1}^n m_i \log[1 + \exp(\x_i \tr \bbeta)]$ term appears in both the $\log \L (\y ; \bbeta)$ and $\pi (\omega_i | \bbeta)$ part, it cancels, thereby justifying the introduction of the $\omega_i$. Finally, we have: 
$$
\log q^*_{\bbeta} (\bbeta) = \bkappa \tr \X \bbeta - \frac{1}{2} \bbeta \tr (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z) \bbeta,
$$
which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma \X \tr \bkappa$.

Next we derive the variational posterior for $\bomega$ in (\ref{eq:qomega}):
\begin{align*}
\log q^*_{\bomega} (\bomega) & = \E_{\bbeta} [\log \pi (\bomega | \bbeta)] = \sum_{i=1}^n \E_{\bbeta} [\log \pi (\omega_i | \bbeta)] \\ & = \sum_{i=1}^n \E_{\bbeta} \( m_i \log \left\{ \cosh \( \frac{\x_i \tr \bbeta}{2}\) \exp\[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i\] \cdot f(\omega_i | m_i, 0) \right\} \) \\ 
& = \sum_{i=1}^n \E_{\bbeta} \[ m_i \log \cosh \( \frac{\x_i \tr \bbeta}{2}\) \] - \frac{1}{2} \sum_{i=1}^n \omega_i \E[ (\x_i \tr \bbeta)^2] + \sum_{i=1}^n \log f(\omega_i | m_i, 0) \\ 
& \propto \sum_{i=1}^n m_i \log \cosh \left\{ \frac{\sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}}{2}\right\} - \frac{1}{2} \sum_{i=1}^n \omega_i \left\{ [\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \right\} \\ 
& \,\,\, + \sum_{i=1}^n \log f(\omega_i | m_i, 0),
\end{align*}
where proportionality is with respect to $\bomega$. If exponentiated, this is a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.

The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated as:
\begin{align*}
\E_{\bbeta} [\log \pi ( \bbeta | \btau)] & = \sum_{j=1}^p \E_{\bbeta_j} [\log \pi(\beta_j | \tau_j )] \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}, \\
\log \pi (\btau) & = \sum_{j=1}^p \log \pi (\tau_j) \propto - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j. 
\end{align*}
Combining the terms we find the log-kernel of the variational posterior as:
\begin{align*}
\log q^*_{\tau} (\btau) & \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}  - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j \\
& = \frac{1}{2} \sum_{j=1}^p \log [(\tau_j - 1)^{-1}] - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2}\tau_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}\),
\end{align*}
where proportionality is with respect to $\btau$. Again applying the change of variables $\psi_j = \tau_j - 1$, we have:
$$
\log q^*_{\psi} (\psi) \propto \frac{1}{2} \sum_{j=1}^p \log \psi_j^{-1} - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2} \psi_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \psi_j^{-1}\),
$$
which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$ random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + \V(\beta_j)]$.

\section{Computational complexity}\label{sec:computationalcomplexity}
Calculation of $\bSigma$ involves the inversion of a $(p \times p)$-dimensional matrix, which occurs at every iteration of the algorithm. If naively done, every inversion is of computational complexity $\mathcal{O}(p^3)$. With $p$ large, this is a serious computational bottleneck of the algorithm, especially with possibly many iterations. However, inspection of the updating equations in (3.9) in the main document, reveals that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the $c_i$ and $\chi_j$. In Section 2.2 of the main document we show that for the EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, and the diagonal of $\bSigma$. Here we provide details on how to compute these with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the computational cost of the algorithm. If necessary, we may calculate the full $\bSigma$ once after the algorithm has converged.

In the following we slightly change notation and let $\diag(\A)$, with $\A$ a $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$ to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$. We start by calculating the complexity of $\diag(\bSigma)$. To this end, let $\H = \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z$. Now, by the Woodbury identity, we may decompose $\diag(\bSigma)$ as:
\begin{subequations}\label{eq:diagsigma}
	\begin{align}
	\diag(\bSigma) ~& = \diag(\H^{-1}) - \diag[\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1}] \\
	& = \diag(\H^{-1}) - [(\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1}) \circ \H^{-1} \X \tr] \cdot \ones_{n \times 1},
	\end{align}
\end{subequations}
where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is a $p \times n$-dimensional matrix of ones. By recognising that $\H$ is a diagonal matrix, we can post-multiply it by a $p \times n$-dimensional matrix with computational complexity $\mathcal{O}(np)$, which leads to a total complexity of $\mathcal{O}(n^2p)$ for $\diag(\bSigma)$. Next we consider the calculation of $\bmu$. Again using the Woodbury identity, we may write $\bmu$ as:
\begin{align}\label{eq:mu}
\bmu = \bSigma \X \tr \bkappa = [\H^{-1} \X \tr - \H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1} \X \tr] \bkappa,
\end{align}
which is, by a similar argument as before, again an operation of complexity $\mathcal{O}(n^2p)$.

Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$ as:
\begin{align*}
\c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
& = \sqrt{\diag \[\X \bSigma \X \tr\] +  \diag \[\X \bmu \bmu \tr \X \tr\]} \\
& = \sqrt{ [(\X \bSigma) \circ \X ] \cdot \ones_{p \times 1} + (\X \bmu) \circ (\X \bmu)}.
\end{align*}
The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$, Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}(n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.

The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p \end{bmatrix} \tr$ are rewritten as:
\begin{subequations}\label{eq:chicomp}
	\begin{align}
	\bm{\chi} & = \begin{bmatrix} \lambda'_{g(1)} \lambda_2 (\bSigma_{11} + \bmu_1^2) & \cdots & \lambda'_{g(p)} \lambda_2 (\bSigma_{pp} + \bmu_p^2) \end{bmatrix} \tr\\
	& = \lambda_2 \bLambda' \[ \diag (\bSigma ) + \diag (\bmu \bmu \tr )\] \\
	& = \lambda_2 \blambda' \circ \[ \diag (\bSigma) + \bmu \circ \bmu \].
	\end{align}
\end{subequations}
 By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we see that the computation of $\bm{\chi}$ consists of one matrix diagonal of $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two Hadamard products of $\mathcal{O}(p)$. Computation of $\bm{\chi}$ is therefore of complexity of $\mathcal{O}(n^2p)$.

\section{Variational Bayes expected joint log likelihood}\label{sec:expectedloglikelihood}
Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then the variational Bayes expected joint log likelihood is derived as follows:
\begin{align*}
\E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ] &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}] \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}(\beta_j | \tau_j) | \blambda'^{(k)}] \\
& \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\ 
& \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
\end{align*}
with the coefficients $d_g^{(k)}$ calculated as:
\begin{align*}
d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\
& = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} ( \beta_j | \blambda'^{(k)} ) \] \\
& = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
\end{align*}

\section{Full estimation procedure}\label{sec:fullestimationprocedure}
The expectation and maximisation steps introduced in the Main Document result in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
\begin{algorithm}
	\caption{Group-regularized empirical Bayes elastic net}\label{alg:summaryalg}
	\begin{algorithmic}[h]
		\Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
		\Ensure $\lambda, \blambda', \bSigma, \bmu$
		\State{Estimate $\lambda$ by cross-validation of the regular elastic net model}
		\While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| > \epsilon_1$}
		\While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
		\State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ for $j=1, \dots, p$ using (3.9) in the Main Document}
		\EndWhile
		\State{Update $\blambda'$ by (3.10) in the Main Document}
		\EndWhile
	\end{algorithmic}
\end{algorithm}

\section{Extensions}\label{sec:extensions}
\subsection{Unpenalized covariates}
Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix $\X$ with a column of ones. Penalization of such an intercept parameter is not desirable \cite[]{Hastie_Elements_2009}. Additionally, it is often desirable to include unpenalized covariates in the model. Common examples of such  covariates in clinical research are patient characteristics, such as age, BMI, and sex. In the following we assume the intercept to be included in the unpenalized covariates.

To include unpenalized covariates, we divide the model matrix into two parts $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$ unpenalized variables and $\X_r$ are the $r$ penalized variables. Let $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with $u$ zero columns and $u$ zero rows. Then we have for the current estimate of the covariance matrix $\bSigma$:
\begin{align}\label{eq:unpencov}
\bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
& = 
\begin{bmatrix} 
\X_u \tr \bOmega \X_u & \X_u \tr \bOmega \X_r \\
\X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
\end{bmatrix}^{-1}.
\end{align}
With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension $p \times p$. Inversion of this matrix is done efficiently by applying the Woodbury identity.

\subsection{Monotonicity of the penalty parameters}
Enforcing monotonicity of the penalty multipliers is desirable in some settings. An example of such a setting is if we have $p$-values from a previous, related study available. A $p$-values-based partitioning of the variables may, \textit{a priori}, be expected to render penalty multipliers that increase monotonically with $p$-value. That is, larger $p$-values may be expected to yield at least as large penalty multipliers as variables with smaller $p$-values. We propose to include this \textit{a priori} assumption by requiring the penalty multipliers to increase monotonically with $p$-value, thereby also stabilizing their estimates.

A natural way of enforcing monotonicity is to extend the constraint in (10) in the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$. The problem is still convex and may be numerically solved. From experience, however, we note that in combination with this constrained optimisation, the EM algorithm described in Section~3 in the Main Document often converges to a local optimum close to the initialisation. We therefore enforce monotonicity through a \textit{post hoc} isotonic regression on the penalty multipliers after every optimisation step.

\subsection{Multiple partitions}
In many cases the features may be partitioned in more than one way. For example, we may have both information on annotation and $p$-values from a previous study available. A naive way of incorporating multiple partitions is to cross-tabulate the partitions and create a separate group for every combination. This poses two problems: (i) The number of penalty parameters increases exponentially with the number of partitions and (ii) some of these combinations may contain only few features, so that the estimation procedure becomes unstable. We propose to stabilise the procedure and keep the number of parameters to estimate manageable by modelling the penalty parameters multiplicatively.

We describe our implementation here for two partitions of the features. To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively. Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively. In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively. The generalised frequentist elastic net estimator and corresponding conditional prior are now:
\begin{subequations}\label{eq:multpart}
	\begin{align}
	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
	\end{align}
\end{subequations}
After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
\begin{subequations}\label{eq:estmultpart}
	\begin{alignat}{2}
	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
	\end{alignat}
\end{subequations}
with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine. This method naturally generalises to more than two partitions of the data.

In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers in a multiplicative way, mainly for convenience. First, there are computational reasons: The separation of the square root in the $L_1$-norm penalty term into two terms facilitates the numerical estimation. Additionally, multiplicative modelling of the partitions allows for more flexibility in the estimates. To see this consider the obvious alternative: additive partition-specific multipliers. Since the penalty multipliers are strictly positive, the constraint on the geometric mean makes one large additive penalty estimate difficult to compensate with a smaller estimate, thereby impairing flexibility of the model. Large multiplicative penalty multipliers are easier to compensate for by smaller penalty multipliers, rendering this the less restrictive option for modelling the partitions.

\section{Simulations}
The quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960}. This kappa is measures similarity between two sets of categorical values. One set contains the feature selection indicators: $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$. We calculate Cohen's $\kappa$ as:
	\begin{align*}
	\kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}\sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
	\end{align*}
	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and expected correctly identified features, respectively. A positive or negative kappa indicate better and worse feature selection than expected by chance, respectively. In addition, we evaluate estimation accuracy by mean squared error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.

\subsection{Scenario 1}
	In this sparse scenario we assume differential signal between 5 groups of features. The average signal in a group increases by a factor 1.6 with group number:
	\begin{align*}
	\bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & \beta_{50}^1 \end{matrix}}_{50}} & \cdots & \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr, \\
	& \, \\
	\beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots, 5  \text{, } j=1, \dots, 50,
	\end{align*}
	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. The model parameters clearly do not follow the relatively heavy-tailed elastic net distribution. Because of this mismatch between the assumed and true distributions, we expect poor performance by \texttt{gren}. Furthermore, we introduce correlation between the features by having covariance elements $\bSigma_{ij} = 0.5^{|i-j|}$.

	\subsection{Scenario 2}
	The second scenario contains many groups of features and is geared towards the non-adaptive group-regularized methods. These methods are able to incorporate the grouping of the features, but were originally designed for small interpretable groups, such as dummies of a categorical variable. \texttt{gren} on the other hand, relies on large groups of features to accurately estimate group-specific penalty weights. The estimation of many penalty weights in small groups is inefficient and computationally challenging. We therefore expect \texttt{gren} to under-perform compared to the other methods.

	For this scenario we follow the second simulation example in \cite{meier_group_2008}. They set $n=100$, $G=250$, and $p=1000$, such that we have 4 features per group. The first 10 groups are active and those 40 model parameters are set to 0.2. The predictors are simulated from a $p$-dimensional Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.

	\subsection{Scenario 3}
	A third scenario is modelled by considering a situation in which there is no differential signal between the groups, but the correlation within groups is very high. We conjecture that this leads to overestimation of effect within the groups by \texttt{gren} and consequently, performance will suffer.

	To that end we sample the predictors in $G=5$ separate blocks of 200 variables, corresponding to the partitioning. The variances of the features are set to one, while all covariances are 0.7. The 1000 model parameters are simulated in two steps. First, we draw 1000 parameters from the elastic net distribution, parametrised as in \cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$. The second step consists of setting the (in absolute value) smallest 100 parameters in a group to zero. This results in a total of 500 zero and 500 non-zero parameters, evenly distributed over the five groups.

	\subsection{Scenario 4}
	This scenario is similar to Scenario 3, except each groups penalty parameter is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the covariance matrix of the predictors is block diagonal, with block sizes $25 \times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$. We expect \texttt{gren} to pick up this differential signal and outperform the other methods.

	\subsection{Scenario 5}
	In the last scenario we create a sparse true model that contains some differential signal. We impl

\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
\subsection{Partitioning based on differential expression}\label{sec:mirna}
The data set is from a deep sequencing analysis on microRNA \cite[]{neerincx_combination_2018}. The study was done in 88 treated colorectal cancer patients, with the aim of classifying treatment response, coded as either non-progressive/remission (70 patients) or progressive (18 patients). After pre-processing and normalisation, 2114 microRNAs remained. In addition to the 2114 microRNAs we incorporated 4 unpenalized clinical covariates into the analysis: prior use of adjuvant therapy (binary), the type of systemic treatment regimen (ternary), age, and primary tumor differentiation (binary).

In a preliminary experiment on different subjects, the microRNA expression levels of metastatic colorectal tumour tissue were compared to normal non-colorectal tissue and primary colorectal tumour tissue was compared to primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 221 microRNAs that were differentially expressed in both comparisons ($\text{FDR} \leq 0.05$), versus 1893 not differentially expressed microRNAs. We expect that incorporation of this partitioning enhances therapy response classification, because tumor-specific miRNAs might be more relevant than non-specific ones. In addition, we divided the differentially expressed microRNAs even further into 127 highly differentially expressed microRNAs ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs ($0.001 < \text{FDR} \leq 0.05$).  We will refer to this second partitioning as the three-group setting, as opposed to the first, the two-group setting.

We compared \texttt{gren} to ridge and elastic net regression, and \texttt{GRridge}. For the elastic net methods (including \texttt{gren}) we set $\alpha \in \{ 0.05, 0.5, 0.95\}$. The estimated penalty multipliers are according to expectation: in all group-regularized methods, the 221 differentially expressed microRNAs receive the smallest penalty (Figure \ref{fig:col_bar}a). In the three-group setting the pattern is again as expected (Figure \ref{fig:col_bar}b): the highly expressed group, receives the lowest penalty, followed by the medium expressed group, while the non-expressed group receives the strongest penalty. \texttt{GRridge} is not able to distinguish between the medium and non-differentially expressed groups of microRNAs.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
	\label{fig:col_bar}
\end{figure}
Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}. All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS. In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC. This is especially true for the larger models. \texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models. A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly. This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation. In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used. Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting. Note that models with few features are often desirable for clinical implementation.
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
	\label{fig:col_per}
\end{figure}

\subsection{Random groups}
Considering that the features may be partitioned into many groups, with one parameter per group, we have to be aware of overfitting risks. We investigated this using the data introduced in Section \ref{sec:mirna}, randomly dividing the features into three groups. We fixed the group sizes to the group sizes used in the three-group setting in Section \ref{sec:mirna}. Under this random partitioning of the features, we expect all penalty multipliers to be estimated as one if overfitting does not occur.

We compared the estimated multipliers to the estimates by \texttt{GRridge}. Since these results depend on one specific randomisation of the groups, we repeated the procedure 100 times and present the results in Figure \ref{fig:random1}. From this figure we see that the estimates for \texttt{gren} are close to one. The estimates by \texttt{GRridge} show much more variation. Additionally, \texttt{GRridge} gives slightly biased estimates: the median penalty parameter estimates are 1.16, 1, and 0.85 for the three groups. In contrast, the median estimates of \texttt{gren} are 1, 1, and 0.99. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\linewidth]{grEBEN_mirseq_Maarten_res6_boxplot.png}
	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 3 random groups in colorectal miRNA data.}
	\label{fig:random1}
\end{figure}

We repeated the simulation with ten evenly sized groups. Figure \ref{fig:random2} presents the estimated penalty multipliers for 100 random splits of the microRNA features into 10 evenly sized groups for \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$. From the Figure we see that \texttt{GRridge} shows more variation around one then the other models. In addition, the median penalty multipliers for the three \texttt{gren} models are all between 0.99 and 1.02, while the \texttt{GRridge} median estimates range from 1.04 to 1.09. This indicates that \texttt{GRridge} suffers from a slight bias in the estimation, while the \texttt{gren} estimates appear to be unbiased.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.9\linewidth]{grEBEN_mirseq_Maarten_res8_boxplot.png}
	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 10 random groups in colorectal miRNA data.}
	\label{fig:random2}
\end{figure}

\subsection{Stability of selection}
One common criticism on the elastic net, and especially the lasso, is the instability of feature selection. That is, different data sets yield different sets of selected features. To investigate the stability of feature selection by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA data \cite[]{neerincx_combination_2018}, where we stratified by treatment response. For each of these bootstrap samples we estimated the regular elastic net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected 25 features (including the 5 unpenalized covariates). We calculated the size of all ${50}\choose{2}$ intersections of selected features and present the results in Figure \ref{fig:bootstrap}. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.9\linewidth]{gren_mirseq_Maarten_overlap.png}
	\caption{Overlap between stratified bootstrap samples for the elastic net models with (a) $\alpha=0.05$, (b) $\alpha=0.5$, and (c) $\alpha=0.95$.}
	\label{fig:bootstrap}
\end{figure}
From this figure it can be seen that stability of feature selection in \texttt{gren} is higher than in the regular elastic net for all three $\alpha$'s.

\section{Empirical-variational Bayes for the ridge and lasso}\label{sec:evbridgelasso}
\subsection{Ridge regression}
Ridge regression is a special case of the elastic net, that does not include an $L_1$-norm on the parameters. Its estimator is given by:
$$
\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_2}{2} \norm{\bbeta}_2.
$$
The Bayesian version of ridge regression is regression under a Gaussian prior. The generalised version of this prior is:
$$
\bbeta \sim \prod_{g=1}^G \prod_{j \in \G(g)} \N(0,(\lambda_2 \lambda'_g)^{-1}).
$$
Although the posterior is analytically calculated in the linear case, in the logistic case we must approximate it iteratively. The variational distributions in this case are:
\begin{align*}
q^*_{\bbeta} (\bbeta) & \sim \mathcal{N} (\bmu, \bSigma), \\ 
q^*_{\bomega} (\bomega) & \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i). \\
\end{align*}
The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net and are updated, together with $\bSigma$, until convergence:
$$
\bSigma^{t+1} = (\X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda')^{-1}\text{, with } \bOmega^{(t)} = \diag \[ \frac{m_i}{2 c^{(t)}_i} \tanh \( \frac{c^{(t)}_i}{2} \) \].
$$

Updating the penalty parameters is done by solving the following (convex) constraint optimisation problem:
\begin{align*}
\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
& \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
\end{align*}
with the terms $d^{(k)}_g$ given by:
$$
d^{(k)}_g = \sum_{j \in \G(g)} ((\bmu^{(k)}_j)^2 + \bSigma^{(k)}_{jj}).
$$

\subsection{Lasso regression}
In lasso regression, another special case of the elastic net, we have the following estimator:
$$
\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1.
$$
The Bayesian counterpart is regression under a Laplace prior on the model parameters. The generalised Bayesian prior may decomposed as:
\begin{align*}
\bbeta | \btau & \sim \prod_{j=1}^p \mathcal{N} \left(0,\tau_j \right), \\
\btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \text{Exp} \left( \frac{\sqrt{\lambda'_g}\lambda_1^2}{8} \right).
\end{align*}

The variational distributions are similar as in the elastic net, except that we do not re-parametrise the $\tau_j$:
\begin{align*}
q^*_{\bbeta} (\bbeta) & \sim \mathcal{N} (\bmu, \bSigma), \\ 
q^*_{\bomega} (\bomega) & \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i), \\
q^*_{\btau}(\btau) & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{GIG} \(\frac{1}{2}, \phi_g, \chi_j\).
\end{align*}
The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic net, while the $\phi_g$ are given by:
$$
\phi_g = \lambda_{g} \frac{\lambda_1^2}{4}.
$$
The following parameters are updated, together with the unchanged $c_i$ and $\bmu$, until convergence:
\begin{align*}
\bSigma^{(t + 1)} &= ( \X \tr \bOmega^{(t)} \X + \Z^{(t)})^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\(\sqrt{\frac{\phi_g(j)}{\chi_j^{(t)}}}\), \\
\chi_j^{(t + 1)} &= \bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \text{, for } j=1, \dots, p.
\end{align*}

Updating the penalty parameters is done by solving the following (convex) constraint problem:
\begin{align*}
\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_1^2}{8} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
& \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
\end{align*}
with the terms $d^{(k)}_g$ given by:
$$
d^{(k)}_g = \sum_{j \in \G(g)} \( \frac{1}{\phi_g^{(k)}} + \frac{\chi^{(k)}_j}{\phi_g^{(k)}} \).
$$ 
Note that the $\phi_g$ are updated after every penalty multiplier update iteration as well, in contrast to the elastic net.

\bibliographystyle{author_short3.bst} 
\bibliography{refs}

\end{document}