### 2019-07-29 docs/biostatistics_main.Rnw
just before resubmit

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("code/figures.R")
@

\documentclass[oupdraft]{bio}
\usepackage{xr-hyper}
% \usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, 
% citecolor=citecolor]{hyperref}
\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,
enumerate}

\externaldocument[sm-]{supplement}

% Add history information for the article if required
%\history{Received August 1, 2010;
%	revised October 1, 2010;
%	accepted for publication November 1, 2010}

% objects
\newcommand{\x}{\mathbf{x}}
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\m}{\mathbf{m}}

% functions and operators
\newcommand{\tr}{^{\text{T}}}
\newcommand{\expit}{\text{expit}}
\newcommand{\argmax}{\text{argmax} \,}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\diag}{\text{diag}}
\newcommand{\trace}{\text{tr}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

\begin{document}

	% Title of paper
	\title{Adaptive group-regularized logistic elastic net regression}

	% List of authors, with corresponding author marked by asterisk
	\author{MAGNUS M. M\"UNCH$^{\ast 1,2}$, CAREL F.W. PEETERS$^{1}$, AAD W. VAN
	DER VAART$^{2}$, \\* MARK A. VAN DE WIEL$^{1,3}$\\[4pt]
	% Author addresses
	\textit{1. Department of Epidemiology \& Biostatistics, Amsterdam Public
	Health research institute, Amsterdam University medical centers, PO Box 7057,
	1007 MB Amsterdam, The Netherlands \\*
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom}
	\\[2pt]
	% E-mail address for correspondence
	{m.munch@vumc.nl}}

	% Running headers of paper:
	\markboth%
	% First field is the short list of authors
	{M. M. M\"unch and others}
	% Second field is the short title of the paper
	{Adaptive group-regularized logistic elastic net regression}

	\maketitle

	% Add a footnote for the corresponding author if one has been
	% identified in the author list
	\footnotetext{To whom correspondence should be addressed.}

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features
		is often available. Examples of such external information in omics research
		are: (a) $p$-values from a previous study, and (b) omics annotation. The
		inclusion of this information in
		the analysis may enhance classification performance and feature selection,
		but is not straightforward. We propose a
		group-regularized (logistic) elastic net
		regression method, where each penalty parameter corresponds to a group of
		features based on the external information. The method, termed
		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
		regression to estimate both the model and penalty parameters in an
		approximate empirical-variational Bayes framework. Simulations and
		applications to three cancer genomics studies and one Alzheimer metabolomics
		study show that, if the partitioning of the features is informative, 
		classification performance and feature selection are indeed enhanced.}
		{Empirical Bayes; High-dimensional data; Prediction; Variational Bayes}
	\end{abstract}

	\section{Introduction}\label{sec:introduction}
	In cancer genomics studies one is often faced with relatively small sample 
	sizes as compared to the number of features. Data pooling may alleviate the 
	curse of dimensionality, but does not apply to research settings with a unique
	set up, and cannot integrate other sources of information. However, external
	information on the features (e.g. genes) is often ubiquitously available in 
	the public domain. We aim to use this information to improve sparse 
	prediction. The information may come in as feature groups: e.g. the chromosome
	on which a gene is located (24 groups), relation to a CpG island for 
	methylation probes (around 6 groups), or membership of a known gene signature 
	(2 groups). Alternatively, it may be continuous, such as $p$-values from a an
	external study. We introduce a method that allows to systematically use
	multiple sources of such external information to improve high-dimensional
	prediction.

  Our methodology is motivated by four recent, small $n$ clinical omics 
  studies, discussed in detail in Section 
  \ref{sec:simandapp}. The studies concern 
  treatment response prediction for colorectal cancer patients, based on 
  sequenced microRNAs ($n$ = 88); Lymph node metastasis prediction for oral 
  cancer, using RNAseq data ($n$=133); Cervival cancer diagnostics based on 
  microRNAs ($n$=56) and Alzheimer diagnosis based on metabolomics ($n$=87).
  Several
  sources of information on the features were used including: external 
  $p$-values, correlation with DNA markers, the conservation status of microRNAs 
  and node degree in an estimated molecular network.

  As basic prediction model we use the (logistic) elastic net 
  regression \cite[]{zou_regularization_2005}, which combines the 
  desirable properties of its special cases ridge
  \cite[]{hoerl_ridge_1970} and lasso regression 
  \cite[]{tibshirani_regression_1996}: de-correlation and feature selection. 
  It has been demonstrated that the prediction accuracy of penalised regression
  can improve by the inclusion of prior knowledge on the variables. Available
  methods, however, either handle one source of external information only
  \cite[]{lee_variable_2017,tai_incorporating_2007}, or do not aim for sparsity
  \cite[]{van_de_wiel_better_2016}.

  Like others, we assume the external information to be available as feature 
  groups; for continuous information like $p$-values we propose a simple, 
  data-based discretisation (see Section \ref{sec:external}).  At first sight, 
  such \textit{a priori} grouping of the features suggests the group lasso 
  \cite[]{meier_group_2008} 
  or one of its extensions such as the group smoothly clipped absolute 
  deviations (\texttt{grSCAD}) and the group minimax concave penalty 
  (\texttt{grMCP}) 
  \cite[]{huang_selective_2012}. These methods penalise and select 
  features at the group level. This comes with two limitations: the group lasso 
  (a) selects entire groups instead of single features and (b) does not
  penalise adaptively: all groups are penalised
  equally. Extensions such as the sparse group lasso 
  \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
  Our way to deal with (b) is through differential penalization. That is, each
  group of features receives its own penalty parameter: the group-regularized 
  elastic net (\texttt{gren}). An apparent issue with differential 
  penalization 
  is the estimation of the penalty parameters. Naive estimation may be done by
  cross-validation (CV). However, CV requires re-estimation of the 
  model over a grid, which grows
  exponentially with the number of penalty parameters. Consequently, it quickly
  becomes computationally infeasible. We therefore propose an efficient 
  alternative: empirical-variational Bayes estimation of the penalty parameters,
  which corresponds to hyperparameter estimation in the Bayesian prior 
  framework. Because of the ubiquity of binary outcome data in clinical omics
  research, we focus on the logistic elastic net.

  Recently, \cite{zhang_novel_2019} introduced a method similar to ours as
  it also applies variational Bayes (VB) for feature selection in logistic
  regression. An advantage of our method, however, is the adaptive inclusion of
  external information on the features to aid in prediction and feature 
  selection, as \cite{zhang_novel_2019} do not estimate feature- or 
  group-specific penalty 
  weights. Bayesian versions of support vector machines have been used in
  classification problems as well \cite[]{chakraborty_bayesian_2011},
  but these methods also lack adaptive inclusion of external information on the
  features. In line with the
  above, our proposed method is (a) data-driven: the
  hyperparameters are estimated from the data, (b) adaptive: prior information
  is automatically
  weighted with respect to its informativeness, (c) fast compared to full 
  Bayesian analysis or CV, and (d) easy to use: only the data and
  grouping of the features are required as input.

  The paper is structured as follows. We introduce the model in Section 
  \ref{sec:model}. In Section 
  \ref{sec:external} we shortly discuss possible sources of co-data.
  In Section \ref{sec:estimation} we derive a VB approximation to the
  model introduced in Section \ref{sec:model}
  and use this novel approximation in the empirical Bayes (EB) estimation of 
  multiple, group-specific penalty parameters. In Section \ref{sec:simandapp}
  we compare the method in a simulation study, demonstrate the benefit of the 
  approach for two 
  data sets, and summarise results for two additional data sets. 
  We conclude with a 
  discussion.

	\section{Model}\label{sec:model}
	The outcome variables are assumed to be binary or sums
	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
	\in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the
	responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix}
	x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} 
	\( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is 
	the binomial distribution with number of trials $m$ and probability 
	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
	\exp(\x\tr_i \bbeta)]$. Throughout the rest of the 
	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
	\x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n
	x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.

	Assume we have a partitioning of the features in $G$ groups, such that each
	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
	$g$ for $g=1, \dots, G$ and let $\lambda'_g \in \mathbb{R}_{>0}$ denote a 
	group-specific penalty weight. In a generalised elastic net regression, the
	penalised likelihood is maximised to yield parameter estimates:
	\begin{equation}\label{eq:weightedlik}
	  \hat{\bbeta} := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  
	  \frac{\lambda_1}{2} 
	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
	\end{equation}
	where $\L(\y ; \bbeta)$ denotes the likelihood function of the 
	observed data $\y = \begin{bmatrix} 
	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
	parameters. From (\ref{eq:weightedlik}) we see that the
	$\lambda'_g$'s may be interpreted as penalty multipliers. Note that
	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.

  Throughout the following we assume that the geometric mean of the multipliers,
  weighted by their 
	respective group sizes, is one, such that the average shrinkage of the model
	parameters is determined by $\lambda_1$ and $\lambda_2$. That is,
	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
	(\lambda'_g)^{|\G(g)|}=1$. The multipliers appear in square root form in the
	$L_1$-norm term to ensure that penalisation on the parameter level scales
	with the norm. The $L_1$-norm sets some of the estimates exactly to zero, thus 
	automatically selecting features. The $L_2$-norm ensures collinearity is 
	well-handled.

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see Supplementary Material (SM) Section 2 for 
	details) may be written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

  \section{External information sources}\label{sec:external}
  We describe possible sources of external information in omics
  studies that may provide the feature groups $\G(g)$. Firstly, there are 
  biologically motivated partitioning of features that are
  easily retrieved from online repositories. These are already 
  discretised and may be included in the analysis as-is. Examples are:
  (i) pathway memberships of genes,
  (ii) classes of metabolites (SM Section 12),
  and (iii) conservation status of microRNAs (SM Section 
  13).
  A second type of external information comes in the form of continuous data.
  Examples are:
  (iv) $p$-values or false discovery rates (FDRs) from a 
  different, related study (Section \ref{sec:colorectal}), and
  (v) quality scores of the features (SM Section 12), 
  and (vii) the node degrees of a network estimated on the feature 
  data (SM Section 12).

  \texttt{gren} requires discretised external information, so in the case of 
  continuous external data, some form of discretisation is required. In some
  cases discretisation comes naturally. For example, with external data 
  type (iv) one might consider the `standard' cut-offs 0.05, 0.01,
  and 0.001. If the choice of cut-offs is not straightforward, we propose a 
  a data-driven heuristic, which renders a relatively finer discretisation grid 
  for data dense than for data sparse domains. Then,
  adaptation to external information is more pronounced in high density areas.
  We propose to fit a piecewise linear spline to the empirical cumulative
  distribution function of the external data and automatically select knot
  locations as cut-offs \cite[]{spiriti_knot_2013}. They also provide a 
  data-driven method to choose the
  number of knots, for a given maximum number of knots. The maximum number
  of knots should be chosen such that each group contains enough features for 
  stable estimation of the penalty weights. As a rule of thumb, we advise
  at least 20 features per group.

  In many practical settings, the external information will be incomplete. 
  We suggest to use a separate
  group of features with missing external information. We prefer this solution
  to setting the penalty multipliers for this group to one, because the absence 
  of external information might be informative from the perspective of 
  prediction.

	\section{Estimation}\label{sec:estimation}
	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
	If the penalty parameters are known, estimation of the elastic net
	model parameters
	is feasible with small adjustments of the available algorithms 
	\cite[]{friedman_regularization_2010,zou_regularization_2005}. Determining 
	these penalty parameters, however, is 
	not straightforward.

	In the frequentist elastic net without group-wise penalisation, two main 
	strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by 
	CV over a two-dimensional grid of values \cite[]{waldron_optimized_2011} or 
	(ii) re-parametrise 
	the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 
	\lambda_2 + \lambda_1}$ and $\lambda = 2 \lambda_2 + \lambda_1$, fix the 
	proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global 
	penalty parameter $\lambda$ \cite[]{friedman_regularization_2010}. In the
	generalised elastic net setting, 
	strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, 
	respectively. $K$-fold CV over $D$ values then results in $K 
	\cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate. Typically $K$ is 
	set to 5, 10, or to the number of samples $n$, while $D$ is in the order of 
	$100$, so that even for small $G$, the number of models to estimate is 
	very large.

	In the Bayesian framework, estimation of penalty parameters may be avoided by 
	the addition of a hyperprior to the model hierarchy. The hyperprior takes the 
	uncertainty in the penalty parameters into account by integrating over them. 
	This approach introduces two issues. Firstly, the choice of hyperprior is not 
	straightforward. Many authors suggest a hyperprior from the gamma family of 
	distributions 
	\cite[]{alhamzawi_bayesian_2018,
	kyung_penalized_2010}, 
	but the precise parametrisation of this gamma prior is not so obvious. 
	Secondly, the correspondence between the  Bayesian and 
	frequentist elastic net is lost. This 
	correspondence may be exploited through the automatic feature selection 
	property of the frequentist elastic net. Endowing the penalty parameters with 
	a hyperprior obstructs their point estimation and, consequently, impedes 
	automatic feature selection. Therefore, to circumvent the problem of 
	hyperprior choice and allow for feature selection by the frequentist elastic 
	net, we propose to estimate the penalty parameters by EB.

	The most formal form of EB is maximisation of
	the marginal likelihood with respect to the hyperparameters. The resulting 
	hyperparameter estimates are then plugged into the prior. The marginal 
	likelihood is often introduced as a measure of model evidence given the 
	observed data and is computed by integrating the product of likelihood and 
	prior with respect to the model parameters. In the case of the elastic net 
	introduced in (\ref{eq:enmodel2}) EB comes down to finding:
	\begin{equation}\label{eq:mmlestimate}
	  \hat{\blambda} := 
	  \underset{\blambda}{\argmax} p_{\blambda}(\y) = 
	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} 
	  \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, 
	  d\bbeta d\btau. 
	\end{equation}
	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
	elastic net. In the omics setting the integrals are also high-dimensional, in 
	which case numerical and Monte Carlo approximation methods become tedious and 
	computationally expensive. Moreover, Laplace approximation is known to suffer 
	from low accuracy in many high-dimensional settings 
	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
	is described that estimates the hyperparameters. This EM algorithm iteratively
	maximises the expected joint log likelihood, such that the sequence:
	\begin{equation}\label{eq:ebEM}
	  \blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
	  \log \[ \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) 
	  \pi_{\blambda}(\btau) \] | \blambda^{(k)} \]
	\end{equation}
	converges to a local maximum of the marginal likelihood. The difficulty herein
	is in the calculation of the expected joint log likelihood. 
	\cite{casella_empirical_2001} suggests to approximate the expectation by its 
	Monte Carlo expectation. Although elegant and simple, this method requires a 
	converged MCMC sample from the posterior for every iteration: a 
	computationally intensive procedure. \cite{roy_selection_2017} introduce
	generalised importance sampling for the Bayesian elastic net, such that we
	just need a limited, pre-specified number of MCMC chains. However, this still
	requires several converged chains, so is not feasible in many high-dimensional
	omics settings. We propose to tackle this problem by approximating the
	expectation in (\ref{eq:ebEM}) using VB.

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
	It has successfully been applied in a wide range of applications, including 
	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
	reconstruction \cite[]{leday_gene_2017}. In VB, the posterior 
	is approximated by a tractable form and estimated by optimizing a lower bound 
	on the marginal likelihood of this model (see SM Section 
	3 for the 
	lower bound of the proposed model). For an extensive introduction and concise 
	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}.

	To simplify the computations of our VB approximation, we follow
	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
	$y_i$ and P\'{o}lya-Gamma distributed (see SM Section 
	4 for more 
	details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	  \bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our VB approximation to the posterior distribution of 
	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
	parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 
	\end{bmatrix} = \begin{bmatrix} \bomega & \bbeta & \btau \end{bmatrix}$, 
	calculus of variations renders the optimal distributions 
	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
	Kullback-Leibler divergence of the posterior to the approximate distribution 
	\cite[]{beal_variational_2003}. The approximation in 
	(\ref{eq:varbayesapproximation})
	renders both the posterior parameter calculations and the expected joint log 
	likelihood in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section 5 
	for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by (13) in SM Section 
	6 until convergence. 
	Naive calculation of the variational
	parameters is computationally expensive. In SM Section 
	7 we show that 
	informed calculation results in a significant reduction of 
	computational complexity.

	\subsection{Empirical-variational Bayes}\label{sec:empvarbayes}
	VB was shown to underestimate the posterior variance of the 
	parameters, both numerically and theoretically, in several settings 
	\cite[]{rue_approximate_2009,
	wang_inadequacy_2005}. This coincides with our experience that the global 
	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated,
	because they are inversely related to the posterior variances of the 
	$\beta_j$. To 
	prevent overestimation we use the parametrisation of
	\cite{friedman_regularization_2010} as discussed in Section 
	\ref{sec:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
	CV of the regular elastic net model, such that the overall 
	penalisation is determined by CV of only $\lambda$. By combining
	CV of the global penalty parameter $\lambda$ with EB estimation 
	of the penalty multipliers $\blambda' = \begin{bmatrix} 
	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is 
	robust to underestimation of the VB posterior variances. 
	For $\alpha$, \cite{hastie_glmnet_2016} recommend
	to either fix it \textit{a priori}, or compare results for several 
	choices of $\alpha$. We use the latter.

	To estimate the penalty multipliers, the intractable posterior expectation
	in (\ref{eq:ebEM}) is approximated using the VB posterior:
	\begin{equation}\label{eq:approximatedloglikelihood}
	  \E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | 
	  \blambda'^{(k)} \]
	  = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
	  \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
	\end{equation}
	where $C$ is constant in $\blambda'$ (see SM Section 
	6 for the full 
	derivation and the $d^{(k)}_g$ terms). An estimate
	of the new penalty multipliers is now given by 
	(14) in SM Section 6. 
	Although the solution to
	(14) in the SM is not available in
	closed form, this convex problem is easily solved by a numerical optimisation 
	routine. The full estimation procedure is summarised in Algorithm 
	\ref{alg:summaryalg}.
	\begin{algorithm}[h!]
    \caption{Group-regularized empirical Bayes elastic net}
    \label{alg:summaryalg}
	  \begin{algorithmic}[h!]
		  \Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
		  \State{Estimate $\lambda$ by CV of the regular elastic net
		  model}
		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
		  \epsilon_1$} 
		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
		  for $j=1, \dots, p$ using (13) in the SM}
		  \EndWhile 
		  \State{Update $\blambda'$ by (14) in the SM}
		  \EndWhile
	  \end{algorithmic}
  \end{algorithm}

	\subsection{Feature selection}\label{sec:featureselection}
	Feature selection is often desirable in high-dimensional prediction problems. 
	For example, biomarker selection may lead 
	to a large reduction in costs by supporting targeted assays. Bayesian feature
	selection is often done by inspection of posterior credible intervals. 
	However, the Bayesian lasso's 
	(a special case of the elastic net) credible intervals  are known to suffer 
	from low frequentist coverage \cite[]{castillo_bayesian_2015}. 
	We therefore propose to select features in the frequentist paradigm.

	Frequentist 
	feature selection is trivial after estimation of the penalty multipliers. We 
	therefore simply plug the estimated penalty parameters into some frequentist
	elastic net
	algorithm that allows for differential penalization. In our package 
	\texttt{gren}, we involve the \texttt{R}-package \texttt{glmnet} 
	\cite[]{friedman_regularization_2010}, which automatically selects features.
	Furthermore, to select a specific number
	of features, we simply adjust the global $\lambda$ to render the desired
	number.

  \section{Simulations and applications}\label{sec:simandapp}
	\subsection{Simulations}\label{sec:simulations}
	We conducted a 
	simulation study in which we compared \texttt{gren} to the regular elastic 
	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016},     
	composite mimimax concave penalty (\texttt{cMCP}) 
	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
	\texttt{gren} in the sense that it estimates group-specific penalty 
	multipliers. The two main differences with \texttt{gren} are (i) the absence 
	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
	are extensions of the group lasso and not adaptive on the group level. 
	However, in contrast to the original group lasso,
	they select single features, instead of complete groups.

	We simulated data according to five different scenarios:
	(i) differential signal between the groups and uniformly distributed model
	parameters; (ii) a large number of small groups of features; (iii) no 
	differential signal between the groups, but strong correlations within groups
	of features; (iv) differential signal between the groups and 
	heavy-tailed distributed model parameters; and (v) a very sparse setting with 
	no signal in some of the groups.

	In all scenarios the $y_i$ are sampled from the logistic model introduced in
	Section \ref{sec:model}, where the $\x_i \tr$ are multivariate Gaussian and
	$\bbeta$ is scenario-dependent. Area under the receiver operator curve (AUC)
	and Brier skill score, averaged over 100 repeats, were used to evaluate 
	performance for models trained on $n=100$ samples and $p \approx 1000$ 
	features. Full descriptions of the scenarios and 
	corresponding results are given in SM Section 
	9. Here, we summarise the results.

	In terms of Brier skill score, the ridge methods generally outperform the 
	elastic net methods, which in turn outperform the group lasso methods.
	In Scenario (i)
	\texttt{gren} and \texttt{GRridge} outperform the other methods in terms
	of AUC.
	In scenario (ii) the AUC follows our expectation: \texttt{gren}, 
	and to
	a lesser extent \texttt{GRridge} underperform due to overfitting. 
	The regular elastic net outperforms the group lasso methods.
	In Scenario (iii), \texttt{gren} and to a lesser extent the
	regular elastic net suffer from the high correlations. The regular elastic net 
	outperforms \texttt{gren}, which is an indication that high correlations
	impair penalty parameter estimation. Scenario (iv) follows the 
	expected pattern, with \texttt{GRridge} and
	\texttt{gren} outperforming their respective non group-regularized 
	counterparts, as well as the sparse group lasso extensions.
	In Scenario (v), \texttt{gren} outperforms all
	other methods, which is an indication that \texttt{gren} is able to pick up 
	the sparse differential signal.

	\subsection{Application to microRNAs in colorectal cancer}
	\label{sec:colorectal}
  We investigated the performance of \texttt{gren} on data from a 
  microRNA sequencing study
  \cite[]{neerincx_combination_2018}. The aim of the study was to predict 
  treatment response in 88 colorectal cancer patients, coded as either
  non-progressive/remission (70 patients) or progressive 
  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
  remained. Four unpenalized clinical
  covariates were included in the analysis: prior use of adjuvant therapy 
  (binary), type of systemic treatment regimen (ternary), age, and primary 
  tumor differentiation (binary).

  In a preliminary experiment on different subjects, the microRNA expression 
  levels of primary and metastatic colorectal tumour tissues were compared to 
  their normal tissue counterparts \cite[]{neerincx_mir_2015}. The two resulting
  FDRs were combined through the harmonic mean \cite[]{wilson_harmonic_2019} 
  and discretised using the method
  described in Section \ref{sec:external}. This yielded four groups of features:
  (i) $\text{FDR} \leq 0.0001$, (ii) $0.0001 < \text{FDR} \leq 0.0186$,
  (iii) $0.0186 < \text{FDR}$, and (iv) missing FDR. We expect that 
  incorporation of this partitioning enhances therapy response 
  classification, because tumor-specific microRNAs are likely to be more 
  relevant than non-specific ones.

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  random forest \cite[]{breiman_random_2001},
  elastic net, sparse group lasso (\texttt{SGL}) by 
  \cite{simon_sparse-group_2013}, \texttt{cMCP}, and
  \texttt{gel}. Of the latter 
  three methods, we only present the best performing one, \texttt{cMCP}, here.
  The results for \texttt{SGL} and \texttt{gel} are presented in SM Section 
  10. For the methods with a
  tuning parameter $\alpha$, we show the best performing $\alpha$ here and refer
  the reader to SM Section 10 for the remaining $\alpha$'s.

  To estimate performance, we split the data into 61 training and 27 test 
  samples, stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present AUC for a
  range of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. Brier skill 
  scores are presented in SM Section 10. 
  In addition, we investigated the sensitivity of
  the multiplier estimation in SM Section 
  10.

  The estimated penalty multipliers are according to expectation: the small
  FDR group receives the lowest penalty, followed by the medium and large
  FDR groups. The missing FDR group receives the strongest penalty, thereby
  confirming that absence of 
  information is informative itself. We observe that
  \texttt{gren} outperforms the other methods for smaller 
  models. Selection of larger models is impaired by the large
  $\alpha$, a property inherent to $L_1$-norm penalization. With a smaller
  $\alpha$ larger models are possible (see SM Section 10). 
  The performance of \texttt{cMCP} is somewhat unstable. This unstable 
  estimation is an issue in all investigated group lasso extensions. 
  Overall, the inclusion of the extra information on the 
  features benefits predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit only slightly for \texttt{GRridge}. The random forest performs worse 
  here.

  \subsection{Application to RNAseq in oral cancer}\label{sec:oral}
  The aim of this second study is to predict lymph node metastasis in
  oral cancer patients using sequenced RNAs from TCGA 
  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
  3096 transformed and normalised TCGA RNASeqv2 gene expression values 
  for 133 HPV-negative oral tumours. Of the corresponding patients, 76 suffered 
  from lymph node metastasis, while 57 
  did not. For a thorough introduction of these data, see 
  \cite{te_beest_improved_2017}.

  We considered two sources of external feature information: (a) the 
  cis-correlations between the RNASeqv2 data and TCGA DNA copy numbers on 
  the same patients, quantified by Kendall's $\tau$ and binned into five groups
  using the rule from Section \ref{sec:external}.
  In addition, we used (b) $p$-values from an independent microarray data set
  described in \cite{mes_prognostic_2017}, again binned into five groups. 
  We expect features 
  with a large positive Kendall's $\tau$ to be more important in 
  metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
  features with low external $p$-values to be more important.

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, since the feature information
  consists of two partitions with overlapping groups, we used extensions of
  \texttt{cMCP} and \texttt{gel} that allow for
  overlapping groups \cite[]{zeng_overlapping_2016}. In \texttt{SGL} we
  cross-tabulated the two partitions to create one grouping of the features.
  We estimated AUC on another independent validation set of 97 
  patients \cite[]{mes_prognostic_2017}, containing microarray features, 
  normalised to account
  for a scale difference with the RNAseq data. We present estimated
  AUC on this validation set for \texttt{GRridge}, ridge, random forest,
  \texttt{gren}, 
  elastic net, and the best performing group lasso extension,
  \texttt{cMCP}, together with the estimated penalty multipliers in Figure
  \ref{fig:barplot_lines_rnaseq_oral_cancer_metastasis_res1_auc}. For the 
  methods with
  an $\alpha$ parameter, we pick the best performing one.
  Results for \texttt{SGL}, \texttt{gel}, other values of $\alpha$, and the 
  Brier skill scores are presented in SM Section 11.

  The estimated penalty multipliers follow the expected pattern: 
  larger cis-correlations receive smaller penalties and smaller $p$-values are
  penalized less. The AUC of
  \texttt{gren} is slighty better than the other methods for a range of model
  sizes. In this example, ridge, \texttt{GRridge}, and random forest
  perform almost 
  identical, while \texttt{gren} outperforms the regular elastic net.

  \subsection{Additional applications}\label{sec:applications}
  Two additional applications are presented in SM Sections 
  12 and 13. The first one is concerned
  with diagnosis of 
  Alzheimer's from 230 metabolites' expression levels in 174 subjects. We 
  included several sources of extra information on the metabolites. In this 
  example, \texttt{gren} performs worse than the group lasso methods for larger
  models. This is 
  due to the smaller number of features to learn the penalty parameters from. 
  Additionally, many metabolites are strongly negatively correlated, which
  further impairs penalty parameter estimation.

  In the second application, the aim is to diagnose 56 women with cervical 
  lesions using 2576 sequenced microRNAs. We include a grouping of the features 
  to enhance predictive performance and 
  feature selection. In this example, \texttt{gren} outperforms the 
  regular elastic net with respect to predictive performance for a range of 
  model sizes. Random forest is competitive to \texttt{gren}, but requires all
  features.

	\section{Discussion}\label{sec:discussion}
	In a taxonomy of Bayesian methods, \texttt{gren} may be considered a 
	local shrinkage model, as opposed to the global-local shrinkage priors that 
	\cite{polson_local_2012} discuss. They characterise 
	certain desirable properties of these global-local shrinkage priors in high 
	dimensions, which, for example, the horseshoe possesses 
	\cite[]{carvalho_horseshoe_2010}. In our case, global 
	shrinkage would imply adding another hyperprior for the global $\lambda_1$ 
	and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters. We argue however,
	that if the groups are informative, the EB estimation of the 
	(semi-) global shrinkage parameters $\lambda'_g$ may be more beneficial than 
	full Bayes shrinkage of the global penalty parameters, because the latter does
	not use any known structure to model variability in the hyperparameters. 
	Nonetheless, an interesting direction of future research is the extension of 
	the group-regularized elastic net to a group-regularized horseshoe model, 
	since the horseshoe has been shown to handle sparsity well and render better
	coverage of credibility intervals than lasso-type priors 
	\cite[]{van_der_pas_horseshoe_2014}.

	Although our method is essentially a reweighted elastic net and can be
	considered weakly adaptive, it is different from 
	the adaptive lasso \cite[]{huang_adaptive_2008,zou_adaptive_2006} and adaptive
	elastic net 
	\cite[]{zou_adaptive_2009} because it adapts to external information
	rather than to the primary data. It also differs in the scale of adaptation: 
	in the adaptive lasso and elastic net the adaptive weights are feature 
	specific, while in our case they are estimated on the group level, rendering
	the adaptation more robust. Both the simulations and the applications
	illustrate that adaptation to 
	external data may be beneficial for prediction and feature selection for a 
	range of marker types (RNAseqs, microRNAs, and metabolites) due to the 
	`borrowing of information' effect: 
	estimates of feature effects that behave similarly are shrunken similarly, 
	yielding overall, better predictions.

	As touched upon in Section \ref{sec:introduction}, an obvious 
	comparison is to the group lasso \cite[]{meier_group_2008} and its extensions.
	Although the group lasso is similar in 
	the sense that it shrinks on the 
	group level, it is built upon an entirely different philosophy: 
	its intended application is to small interpretable groups of features, like, 
	for example, gene pathways. Another difference between 
	\texttt{gren} and the group lasso is the form of the penalty: 
	\texttt{gren} estimates one parameter per group. Once these are estimated
	\texttt{gren} fits a reweigthed elastic net. The group lasso uses one overall
	penalty parameter; 
	it is thereby less flexible in differential shrinkage of the parameters. Our 
	simulations and data applications show that \texttt{gren} is competitive and
	often superior to (extensions of) the group lasso.

	A recent development is to group samples
	rather than features \cite[]{dondelinger_joint_2018} to allow different levels 
	of sparsity across sample groups. This approach does not incorporate feature
	information, and uses CV to estimate the penalty parameters. An
	interesting future line of research would be to combine this approach with 
	\texttt{gren}.

	A common criticism of the lasso (and elastic net) is its instability of 
	feature 
	selection: different data instances lead to different sets of selected
	features. To investigate the stability of selection on the data from Section
	\ref{sec:colorectal} we created bootstrap samples from 
	the original data and calculated the sizes of the selected feature set 
	intersections. Compared to the regular elastic net, the inclusion of extra 
	information
	increases the selection stability (see SM Section 10). 
	In addition, we investigated penalty multiplier estimation 
	on 100 random partitionings of the features
  (SM Section 10). We found that the estimated penalty 
  parameters tend to cluster around one, as desired with random groups.

	A possible weak point of the proposed method is the double EM loop, which may
	end up in a local optimum, depending on the starting values of the algorithm.
	Reasonable starting values, e.g. obtained by applying \texttt{GRridge}, could
	alleviate this issue.
	The default of \texttt{gren}, however, is to simply start from the regular
	elstic net, i.e., penalty multipliers equal to one, and than adapt these. In 
	the applications, we investigated the occurrence of multiple optima, but never
	encountered them. This does not guarantee that local optima do not occur, but 
	it provides some evidence that local optima are not ubiquitous.

	Proper uncertainty quantification in the elastic net is an open problem.
	If uncertainy quantification is required, Gibbs samples from 
  the posterior (with the estimated hyperparameters) could be drawn. However,
  we do not recommend Bayesian uncertainty quantification
  for the Bayesian elastic net due to bad frequentist properties of lasso-like
  posteriors \cite[]{castillo_bayesian_2015} in sparse (omics) settings. Hence,
  selected features should be interpreted with caution, but are nonetheless
  deemed useful in prediction.

  An EM algorithm runs the danger of excessive computation time. In our 
  implementation we have reduced computational time considerably by
  some computational shortcuts (see SM Section 
  7) and implementing some
  parts in \texttt{C++}. To assess computation times we compared \texttt{gren}
  other methods introduced in Sections \ref{sec:simulations}-
  \ref{sec:oral} on a Macbook Pro 2016 running 
  \Sexpr{gsub("_", ".", devtools::session_info()$platform$system)} and present the
  results in SM Section 14. In general we found that 
  \texttt{gren} is similar in times as \texttt{GRridge}, faster than 
  \texttt{SGL}, and slower than \texttt{cMCP}, and \texttt{gel}.

	\section*{Software}
	A (stable) \texttt{R} package is available from 
	\url{https://CRAN.R-project.org/package=gren}.

	\section*{Supplementary Material}
	Supplementary Material is available online from 
	\url{http://biostatistics.oxfordjournals.org}.

	\section*{Reproducible Research}
	All results and documents may be recreated from 
	\url{https://github.com/magnusmunch/gren}.

	\section*{Funding}
	This research has received funding from the European Research Council under 
	ERC Grant Agreement 320637.

	\section*{Acknowledgements}
	We thank the editor and referees for their useful suggestions and comments. 
	\textit{Conflict of Interest}: None declared.

	\bibliographystyle{biorefs}
	\bibliography{refs}

<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@

<<barplot_lines_rnaseq_oral_cancer_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
@

\end{document}